---
title: "Introducing R packages at Rapporteket"
author: "Are Edvardsen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Hoepfully, use of R packages will make both development and system meintenance at Rappporteket.
However, this calls for some changes of both routines and configuration/set-up. In the
following, such changes are logged as they are made. Hence, all entries here should be
reviewed before an approval of this approach.

## User privileges
Intallation of packages must also work automated. In a manual approach operating with
root privileges is common and trivial. However, processes either automated or
initiated by random users or external triggers should always be run with very reduced
privileges. At _Rapporteket_ all R related processes are run by the unprivileged user
_tomcat_. This should the also apply to handling of R packages.

During installation of R packages R checks the privileges of the user who triggers the
installation. If the user has root privileges, the package is installed alongside the rest
of the (default) packages providing system wide access. If the user does not have root 
privileges the package will be installed in the users home directory and will thus only be
available to the current user.

The _tomcat_ user does not even have his own home directory suitable for install of R
packages. Hence, a local R package site library is proposed where _tomcat_ have read
and write access. By default, an R install comes with built in definition of several options,
and at Rapporteket these are:

```r
Sys.getenv("R_LIBS_SITE")
[1] "/usr/local/lib/R/site-library:/usr/local/lib/R/library:/usr/lib64/R/library:/usr/share/R/library"
```
The only one that actually exists is _/usr/lib64/R/library_ and packages installed as
root are all placed here. User _tomcat_ does not have write access to this directory and
cannot install any packages.

To enable user _tomcat_ to install our own packages it is proposed to use on of the
other options, by (as root) creating directories and setting appropriate privileges:

```bash
cd /usr/local/lib
mkdir -p R/site-library
chown -R root:tomcat R
chmod -R ug+rwx R
```
 Now, also user _tomcat_ will be able to install packages.
 
 An alternative approch will be to create a proper HOME for user _tomcat_.





Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
